---
title: "Bayes’ Theorem Flu Test Example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bayes’ Theorem Flu Test Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
library(bayesToolkit)
library(ggplot2)
```

## Overview
Bayes’ Theorem is a foundational concept in probability, statistics, and machine
learning. In healthcare analytics, it is frequently used to update clinical risk
estimates as new patient data becomes available.

This vignette walks through a realistic example using influenza testing data to
compute:

- the probability of having the flu given a positive test result
- a full 2×2 contingency table
- accessible visualizations generated by the bayes() function

## Flu Test Scenario
Public health data provide the following estimates:

- Prevalence of flu in the population: $P(A) = 0.07$
- Sensitivity of the flu test (true positive rate): $P(B \mid A) = 0.92$
- Specificity of the test: $P(B^- \mid A^-) = 0.93$, which implies a false
positive rate of $P(B \mid A^-) = 0.07$

Given this detail, what is the probability of actually having the flu if a patient tests
positive?

## Using the bayes() Function

### Using the Function
```{r}
result <- bayes(
  prevalence = 0.07,
  sensitivity = 0.92,
  false_positive = 0.07
)
```

### Posterior Probability

This value represents: $P(A \mid B)$, which is the probability of having the
flu given a positive test.

```{r}
result$posterior_probability
```

### Contingency Table

This table shows the scaled counts (default population size = 1000):

- True Positives
- False Positives
- False Negatives
- True Negatives

```{r}
result$contingency_table
```

### Visualizations
#### Grouped Bar Chart
```{r, fig.width=6, fig.height=4}
result$barchart
```

#### Heatmap
```{r, fig.width=6, fig.height=4}
result$heatmap
```

#### Posterior Probability Plot
```{r, fig.width=6, fig.height=4}
result$posterior_plot
```

## Hand Calculation
Bayes' Theorem calculations can be performed manually for the validation of
results:

*Given*

- $P(A) = 0.07$
- $P(A^-) = 1 - P(A) = 0.93$
- $P(B \mid A) = 0.92$
- $P(B \mid A^-) = 0.07$

*Compute the probability of a positive test*

- Calculate $P(B)$, which is the probability of a positive test, calculated
using the Law of Total Probability
- Through Bayes' Theorem, compute $P(A \mid B)$, which is the probability of
having the flu given a positive test result

---

## References
- Centers for Disease Control and Prevention. (n.d.). _Flu burden estimates and data visualization tools._ https://www.cdc.gov/flu-burden/php/data-vis/index.htmlLinks to an external site.
- Centers for Disease Control and Prevention. (2024, September 17). _Rapid influenza diagnostic tests._ https://www.cdc.gov/flu/hcp/testing-methods/rapidclin.htmlLinks to an external site.
- Centers for Disease Control and Prevention. Flu burden estimates and data visualization tools.
- Statista Research Department. (2026, April 23). _Total population of the United States from 2015 to 2024, with a forecast until 2031._ Statista. https://www.statista.com/statistics/263762/total-population-of-the-united-states/Links to an external site.